---
title: "BACS HW15"
author: '109090046'
date: "2023-05-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(seminr)
```

## Load Dataset
```{r cars}
security_sem <- read.csv('D:/桌面/NTHU/Courses/BACS/HW15/security_data_sem.csv')
```

## Question 1) Composite Path Models using PLS-PM

### a. Create a PLS path model using SEMinR, with all the following characteristics:

#### i. Measurement model – all constructs are measured as composites:
```{r}
security_mm <- constructs(
  composite('TRUST', multi_items('TRST', 1:4)),
  composite('SEC', multi_items('PSEC', 1:4)),
  composite('REP', multi_items('PREP', 1:4)),
  composite('INV', multi_items('PINV', 1:3)),
  composite('POL', multi_items('PPSS', 1:3)),
  composite('FAML', single_item('FAML1')),
  interaction_term(iv='REP', moderator='POL', method=orthogonal)
)
```

#### ii. Structural Model – paths between constructs as shown in this causal model:
**REP + INV + POL + FAML + (REP✕POL) → SEC → TRUST**
```{r}
security_sm <- relationships(
  paths(from = c("REP","INV","POL","FAML","REP*POL"), to = "SEC"),
  paths(from = "SEC", to = "TRUST")
)
```

```{r}
pls <- estimate_pls(
  data = security_sem,
  measurement_model = security_mm,
  structural_model = security_sm
)

summary(pls)
```

### b. Show us the following results in table or figure formats:

#### i. Plot a figure of the estimated model
```{r}
plot(pls, title="PLS plot")
```

#### ii. Weights and loadings of composites
```{r}
pls_summary <- summary(pls)
# Weight of composites
pls_summary$weights

# Loading of composites
pls_summary$loadings
```

#### iii. Regression coefficients of paths between factors
```{r}
pls_summary$paths
```

#### iv. Bootstrapped path coefficients: t-values, 95% CI
```{r}
boot_pls <- bootstrap_model(pls, nboot=1000)

summary(boot_pls)
```

## Question 2) Common-Factor Models using CB-SEM

### a. Create a common factor model using SEMinR, with the following characteristics:

#### i. Either respecify all the constructs as being `reflective()`, or use the `as.reflective()` function to convert your earlier measurement model to being entirely reflective.
```{r}
ref_mm <- as.reflective(security_mm)
```

##### ii. Use the same structural model as before (you can just reuse it again!)
```{r}
security_sm
```

### b. Show us the following results in table or figure formats
```{r}
sec_cbsem <- estimate_cbsem(
  data = security_sem,
  measurement_model = ref_mm,
  structural_model = security_sm
)

summary(sec_cbsem)
```

#### i. Plot a figure of the estimated model (it will look different from your PLS model!)
```{r}
plot(sec_cbsem, title="Common-Factor Models using CB-SEM")
```

#### ii. Loadings of composites
```{r}
summary_cbsem <- summary(sec_cbsem)
summary_cbsem$loadings
```

#### iii. Regression coefficients of paths between factors, and their p-values
```{r}
summary_cbsem$paths
```